\documentclass[12pt]{article}

\usepackage[margin=1.12in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\C}{\mathbb C}
\newcommand{\Z}{\mathcal Z}
\newcommand{\taua}{\tau_a}
\newcommand{\taub}{\tau_b}
\newcommand{\supp}{\operatorname{supp}}

\setlength{\emergencystretch}{3em}
\hypersetup{
  pdftitle={Relative Entropy Does Not Supply an Omitted Finite-Central Calibration},
  pdfsubject={Relative entropy, finite centers, JLMS, OAQEC, entropy calibration},
  pdfkeywords={relative entropy, finite centers, JLMS, operator algebra quantum error correction, entropy calibration}
}

\title{Relative Entropy Does Not Supply\\
an Omitted Finite-Central Calibration}
\author{}
\date{Live review note of June 20, 2026}

\begin{document}
\maketitle

\begin{abstract}
Relative entropy and JLMS/OAQEC recovery comparisons are statements about the
algebra and state family supplied to them.  This note records the finite-center
boundary.  In a finite central decomposition with trace weights
\(\tau_a(c)=\sum_i a_i c_i\), the density of a central probability vector
\(p\) is \(p_i/a_i\).  Hence
\[
  S_{\tau_a}(p\Vert r)
  =
  \sum_i p_i\log{p_i\over r_i},
\]
so the common trace weights cancel from every pairwise relative entropy on a
fixed retained face.  The absolute entropy still contains the affine central
term \(\sum_i p_i\log a_i\).  Therefore relative-entropy equality does not, by
itself, determine an omitted finite central trace/area calibration or an
absolute entropy-density representative.  The positive exit is an affine or
atomic calibration that fixes this central term on the state family, or a
physical axiom excluding the hidden center.
\end{abstract}

\paragraph{Status.}
This is non-frozen external-review metadata for Team 1.  It is intended as a
small claim that a relative-entropy, OAQEC/JLMS, crossed-product, or
black-hole-entropy reviewer can classify as known, false, missing-hypothesis,
or useful.  It does not modify any frozen Team 1 manuscript, Lean archive,
public mirror manifest, or SHA-256 package.

\section{Primary sources}

The boundary below was checked against Araki's relative entropy for von
Neumann algebras \cite{ArakiRelativeEntropy}, the JLMS relative-entropy
equality \cite{JLMS}, entanglement-wedge reconstruction
\cite{DongHarlowWall}, Harlow's operator-algebra quantum error-correction
formulation of the RT formula \cite{HarlowRTQEC}, and gravitational
crossed-product entropy \cite{WittenCrossed,CPWLargeN}.  These works are used
as positive statements for supplied algebras, traces, and state families.  No
new relative-entropy theorem is claimed here.

\section{Finite central face}

Let
\[
  \Z_m=\ell^\infty(\{1,\ldots,m\})\cong\C^m
\]
and let \(\tau_a(c)=\sum_i a_i c_i\), where \(a_i>0\).  If \(p\) is a central
probability vector, its density with respect to \(\tau_a\) is \(p_i/a_i\).
For two probability vectors \(p,r\) with \(\supp p\subseteq\supp r\),
\[
  S_{\tau_a}(p\Vert r)
  =
  \sum_{i:p_i>0}p_i
  \log{(p_i/a_i)\over(r_i/a_i)}
  =
  \sum_{i:p_i>0}p_i\log{p_i\over r_i}.
  \tag{1}
\]
Thus the trace weights cancel from pairwise relative entropy on the retained
central face.

The absolute entropy relative to the same trace is
\[
  H_{\tau_a}(p)
  =
  -\sum_i p_i\log{p_i\over a_i}
  =
  H(p)+\sum_i p_i\log a_i .
  \tag{2}
\]
The second term is an affine central calibration.

\begin{proposition}[relative entropy leaves a calibration fiber]
\label{prop:calibration-fiber}
Fix a retained central face \(F\subseteq\{1,\ldots,m\}\) and a state family
\(P\) supported on \(F\).  If \(a\) and \(b\) are faithful trace calibrations
on \(F\), then every pairwise relative entropy among states in \(P\) is the
same for \(\tau_a\) and \(\tau_b\).  Consequently pairwise relative entropy on
\(P\) does not determine the trace weights, the affine term
\(\sum_i p_i\log a_i\), or any absolute entropy-density coefficient depending
on that term.
\end{proposition}

\begin{proof}
Equation (1) contains only the probability ratios \(p_i/r_i\).  The weights
\(a_i\) cancel.  Replacing \(a\) by \(b\) therefore leaves all pairwise
relative entropies unchanged.  Equation (2) shows that absolute entropy still
depends on the affine functional \(p\mapsto\sum_i p_i\log a_i\), which has not
been fixed by those relative entropies.
\end{proof}

\begin{proposition}[two-atom witness]
\label{prop:two-atom-witness}
On the two-atom center, let \(a=(1,3)\), \(b=(2,2)\), and
\(p=(1/2,1/2)\).  The pairwise relative entropies on the two-atom probability
simplex are unchanged when \(\tau_a\) is replaced by \(\tau_b\), but
\[
  H_{\tau_a}(p)-H_{\tau_b}(p)=\frac12\log{3\over4}\ne0 .
\]
\end{proposition}

\begin{proof}
The first statement is Proposition~\ref{prop:calibration-fiber}.  For the
second,
\[
  H_{\tau_a}(p)-H_{\tau_b}(p)
  =
  {1\over2}(\log1+\log3-\log2-\log2)
  =
  {1\over2}\log{3\over4}.
\]
\end{proof}

\begin{theorem}[affine central calibration criterion]
\label{thm:affine-calibration}
Let \(P\) be the retained central state family and write
\[
  L_a(p)=\sum_i p_i\log a_i .
\]
An absolute central entropy or area calibration determines the finite-center
entropy term on \(P\) exactly when it determines the affine functional \(L_a\)
on the affine hull of \(P\).  In particular, vertex calibrations on every
central atom recover all \(\log a_i\) up to the common trace-normalization
convention; calibrations on a proper face leave the omitted vertices
unconstrained.
\end{theorem}

\begin{proof}
By (2), all dependence on trace calibration is through the affine functional
\(L_a\).  Knowing absolute entropy or area values on an affinely spanning
subset of \(P\) fixes \(L_a\) on the affine hull of \(P\), and no smaller
calibration can fix it outside its affine span.  Vertex tests read off the
values \(L_a(e_i)=\log a_i\).  If a vertex is omitted, changing only that
coordinate gives another faithful calibration with the same supplied tests and
different absolute entropy for states using the omitted vertex.  A common
rescaling \(a\mapsto\lambda a\) adds the constant \(\log\lambda\) to all
\(L_a(p)\), which is invisible only when the intended convention is
projective.
\end{proof}

\section{JLMS/OAQEC boundary}

JLMS and operator-algebra quantum error correction give powerful positive
statements for the declared algebra and state family.  They become a positive
exit from this Team 1 obstruction exactly when the comparison includes the
central face and its absolute affine calibration.

If the comparison factors through a quotient, selected central character,
declared code algebra, or state family that omits some central vertices, then
relative entropy remains algebra-relative.  It does not certify maximality of
the finite center or determine the omitted trace/area calibration.

\section{Referee question}

\begin{quote}
In the target black-hole/AQFT reconstruction setting, does the invoked
JLMS/OAQEC/crossed-product result supply full central-face affine calibration,
or only relative entropy for the algebra and state family already specified?
\end{quote}

If the answer is the second, Team 1 has no new relative-entropy theorem.  The
surviving claim is only the claim-boundary warning: relative entropy and
recovery data should not be overread as finite-center maximality or as an
absolute finite-center calibration principle.

\begin{thebibliography}{9}

\bibitem{ArakiRelativeEntropy}
H. Araki,
\emph{Relative entropy of states of von Neumann algebras},
Publ. RIMS 11 (1976), 809--833,
\href{https://ems.press/content/serial-article-files/41836}{EMS PDF}.

\bibitem{JLMS}
D. L. Jafferis, A. Lewkowycz, J. Maldacena, and S. J. Suh,
\emph{Relative entropy equals bulk relative entropy},
JHEP 06 (2016), 004,
\href{https://arxiv.org/abs/1512.06431}{arXiv:1512.06431}.

\bibitem{DongHarlowWall}
X. Dong, D. Harlow, and A. C. Wall,
\emph{Reconstruction of Bulk Operators within the Entanglement Wedge in
Gauge-Gravity Duality},
Phys. Rev. Lett. 117 (2016), 021601,
\href{https://arxiv.org/abs/1601.05416}{arXiv:1601.05416}.

\bibitem{HarlowRTQEC}
D. Harlow,
\emph{The Ryu-Takayanagi Formula from Quantum Error Correction},
Commun. Math. Phys. 354 (2017), 865--912,
\href{https://arxiv.org/abs/1607.03901}{arXiv:1607.03901}.

\bibitem{WittenCrossed}
E. Witten,
\emph{Gravity and the Crossed Product},
JHEP 10 (2022), 008,
\href{https://arxiv.org/abs/2112.12828}{arXiv:2112.12828}.

\bibitem{CPWLargeN}
V. Chandrasekaran, G. Penington, and E. Witten,
\emph{Large N algebras and generalized entropy},
JHEP 04 (2023), 009,
\href{https://arxiv.org/abs/2209.10454}{arXiv:2209.10454}.

\end{thebibliography}

\end{document}
